---
title: "locksmith_goals"
author: "tim bender"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview of NC Locksmithing

In North Carolina, performing locksmith services requires a license from the NC Locksmith Licensing Board, obtained by being at least 18, having good moral character, passing an exam, and paying fees. Licensed locksmiths must also take reasonable steps to verify that the customer is authorized to access the property and must carry and present their license upon request. Some exceptions exist, such as employees working under direct supervision or individuals performing only administrative duties.

```{r cars}
# code chunk
```

## NC General Statutes - Chapter 74F
Full language [here](https://www.ncleg.net/EnactedLegislation/Statutes/PDF/ByChapter/Chapter_74F.pdf).


#### § 74F-1. Short title.
This act shall be known as the North Carolina Locksmith Licensing Act. (2001-369, s. 1.)

#### § 74F-2. Purpose.
Locksmiths have the knowledge and tools to bypass or neutralize security devices in vehicles,
homes, and businesses. The laws of this State do not protect citizens from the unscrupulous use and
abuse of this knowledge and these tools by persons who are untrained or have criminal intent.
Therefore, **the licensing of locksmiths is necessary to protect public health, safety, and welfare**.
(2001-369, s. 1.)

#### § 74F-3. Licenses required; violation.
  (a) No person shall perform or offer to perform locksmith services in this State unless the
person has been licensed under the provisions of this Chapter. Every person providing locksmith
services as defined under G.S. 74F-4(5) **to buildings containing medical records, pharmaceutical records, educational records, criminal records, voting records, tax records, legal records, or personnel records, including any person providing locksmith services who is employed by or working for a school, college, university, hospital, company, institution, or government facility shall be licensed** in accordance with the provisions of this Chapter.
  (b) Unless the conduct is covered under some other provision of law providing greater
punishment, a violation of this section is a **Class 1 misdemeanor for the first offense. A second or subsequent offense is a Class I felony**. (2001-369, s. 1; 2013-370, ss. 1, 2.)

#### § 74F-4. Definitions.
The following definitions apply in this Chapter:

  (1) Apprentice. – A person who has been issued an apprenticeship designation by
the Board.

    (1a) Board. – The North Carolina Locksmith Licensing Board.

  (2) Code book. – A compilation, in any form, of key codes and combinations.

  (3) License. – A certificate issued by the Board recognizing the person named
therein as having met the requirements to perform locksmith services as defined
in this Chapter.
(4) Locksmith. – A person who has been issued a license by the Board.
(5) Locksmith services. – Services that include repairing, rebuilding, rekeying,
repinning, servicing, adjusting, or installing locks, mechanical or electronic
locking devices, access control devices, egress control devices, safes, vaults,
and safe-deposit boxes for compensation or other consideration, including
services performed by safe technicians. The definition also includes any method
of bypassing a locking mechanism of any kind, whether in a commercial,
residential, or automotive setting, for compensation.
(6) Locksmith tools. – Any tools that are designed or used to open a mechanical or
electrical locking device in a way other than that which was intended by the
manufacturer. (2001-369, s. 1; 2003-350, ss. 1, 2; 2013-370, s. 3.)

#### § 74F-5. North Carolina Locksmith Licensing Board.

(a) Composition and Terms. – The Board shall consist of nine members who shall serve
staggered terms. Three members shall represent the public. The initial Board members shall be
appointed on or before January 1, 2002, as follows:
(1) The General Assembly, upon the recommendation of the President Pro Tempore
of the Senate, shall appoint three locksmiths, two of whom shall serve terms of
four years and one of whom shall serve a term of three years. At least one of the
locksmiths shall represent a recognized locksmith organization in the State.
(2) The General Assembly, upon the recommendation of the Speaker of the House
of Representatives, shall appoint three locksmiths, one of whom shall serve a
term of four years, one of whom shall serve a term of three years, and one of
whom shall serve a term of two years. At least one of the locksmiths shall
represent a recognized locksmith organization in the State.
(3) The Governor shall appoint three public members, one of whom shall serve a
term of three years and two of whom shall serve terms of two years.
Upon the expiration of the terms of the initial Board members, each member shall be appointed
for a term of three years and shall serve until a successor is appointed. No member may serve more
than two consecutive terms.
(b) Qualifications. – The locksmith members shall have at least five years' experience in
locksmith services and shall be engaged in that business for the duration of their term on the Board.
The locksmith members initially appointed to the Board shall immediately become licensed as
locksmiths by complying with the provisions of this Chapter. Public members of the Board shall
not be trained or experienced in locksmith services, have a financial interest in a locksmith
business, or be the spouse of a person who is so trained or experienced or has such an interest. All
members of the Board shall reside in this State and shall represent various geographical areas of the
State.
(c) Vacancies. – A vacancy shall be filled in the same manner as the original appointment,
except that all unexpired terms in seats appointed by the General Assembly shall be filled in
accordance with G.S. 120-122. Appointees to fill vacancies shall serve the remainder of the
unexpired term and until their successors have been duly appointed and qualified.
(d) Removal. – The Board may remove any of its members for neglect of duty,
incompetence, or unprofessional conduct. A member subject to disciplinary proceedings as a
licensee shall be disqualified from participating in the official business of the Board until the
charges have been resolved.
(e) Compensation. – Each member of the Board shall receive per diem and reimbursement
for travel and subsistence as provided in G.S. 93B-5.
(f) Officers. – The officers of the Board shall be a chair, a vice-chair, and other officers
deemed necessary by the Board to carry out the purposes of this Chapter. All officers shall be
elected annually during the first meeting of the calendar year by the Board for one-year terms and
shall serve until their successors are elected and qualified.
(g) Meetings. – The Board shall hold at least two meetings each year to conduct business
and to review the standards and rules for issuing licenses under this Chapter. The Board shall adopt
rules governing the calling, holding, and conducting of regular and special meetings. A majority of
Board members shall constitute a quorum. (2001-369, s. 1.)

#### § 74F-6. Powers of the Board.
The Board shall have the power and duty to:

  (1) Administer and enforce the provisions of this Chapter.
  (2) Adopt rules as may be necessary to carry out the provisions of this Chapter.
  (3) Examine and determine the qualifications and fitness of applicants for licensure and renewal of licensure.
  (4) Issue, renew, deny, suspend, or revoke licenses or apprenticeship designations and conduct any disciplinary actions authorized by this Chapter.
  (5) Set fees as provided in G.S. 74F-9.
  (6) Establish and approve continuing education requirements for persons licensed under this Chapter.
  (7) Receive and investigate complaints from members of the public.
  (8) Conduct investigations for the purpose of determining whether violations of this Chapter or grounds for disciplining licensees exist.
  (9) Conduct administrative hearings in accordance with Article 3A of Chapter 150B of the General Statutes.
  (10) Maintain a record of all proceedings conducted by the Board and make available to licensees and other concerned parties an annual report of all Board action.
  (11) Maintain a list of the names and addresses of all persons licensed by the Board.
  (12) Employ and fix the compensation of personnel that the Board determines is necessary to carry out the provisions of this Chapter and incur other expenses necessary to perform the duties of the Board.
  (13) Adopt and publish a code of ethics.
  (14) Adopt a seal containing the name of the Board for use on all licenses and official reports issued by the Board.
  (15) Employ an attorney to assist or represent the Board in enforcing this Chapter.
  (16) Request that the State Bureau of Investigation conduct criminal history record checks of applicants for licensure and apprenticeships pursuant to G.S. 143B-1209.26 [G.S. 143B-1209.27].
  (17) Authorize the chair, by majority vote, to issue subpoenas allowing the Board to obtain the records of a person or company offering locksmith services, including an employee of a company, a contractor, or a subcontractor. The records obtained shall include invoices and receipts, specifically any invoices and receipts that pertain to locksmith tools, equipment, or parts. (2001-369, s. 1; 2003-350, ss. 3, 4; 2004-177, s. 1; 2013-370, s. 4; 2014-100, s. 17.1(gg); 2023-134, s. 19F.4(z).)

#### § 74F-7. Qualifications for license.
An applicant shall be licensed as a locksmith if the applicant meets all of the following
qualifications:

  (1) Is of good moral and ethical character, as evidenced in part by a criminal history record check conducted in accordance with G.S. 74F-18.
  (2) Is at least 18 years of age.
  (3) Successfully completes an examination administered by the Board that measures the knowledge and skill of the applicant in locksmith services and the laws applicable to licensed locksmiths.
  (4) Pays the required fee under G.S. 74F-9. (2001-369, s. 1; 2003-350, s. 5.)

#### § 74F-7.1. Apprentices.
(a) An applicant may receive an apprentice designation if the applicant meets all of the
following requirements:

  (1) Is of good moral and ethical character, as evidenced in part by a criminal history record check conducted in accordance with G.S. 74F-18.
  (2) Is at least 18 years of age.
  (3) Pays the required fee under G.S. 74F-9.
    
(b) After an applicant has satisfied the requirements in subsection (a) of this section, the
apprentice may practice as an apprentice locksmith under the supervision of a licensed locksmith.
The Board shall issue each apprentice a colored badge card that identifies the individual as an
apprentice and includes the apprentice designation. The Board shall establish requirements for
apprentice supervision.

(c) An apprentice may have the apprentice designation for no longer than three years. On
or before the three-year period expires, the apprentice shall take the locksmith licensure
examination administered by the Board. If the apprentice fails to take the examination within the
three-year period or fails the examination and does not retake the examination on or before the
three-year period expires, the apprentice shall not receive licensure and shall not be granted
another apprenticeship.

(d) If an apprentice terminates employment with a licensed locksmith under which the
apprentice originally received his or her apprenticeship, the apprentice shall find employment with
another licensed locksmith to maintain the apprentice designation and shall pay a transfer fee
pursuant to G.S. 74F-9.
Each licensed locksmith shall have no more than two apprentices at one time. However, a
licensed locksmith shall have a 90-day grace period within which to accommodate more than two
apprentices if the apprentice is newly hired as a result of a previous termination of employment or
the inability of a licensed locksmith to supervise the apprentice. The licensed locksmith shall
contact the Board once the locksmith terminates the additional apprentice, and the licensed
locksmith shall be legally responsible for the apprentice's work until the Board is notified
otherwise. (2003-350, s. 6.)

#### § 74F-8. Licensure based on experience; licensure of nonresident; reciprocity.
    
  (a) The Board may grant, upon application and payment of proper fees, a license to a person who resides in this State and has at least three years' experience as a licensed locksmith in another state whose standards of competency are substantially equivalent to those provided in this Chapter.
  (b) The Board may grant, upon application and payment of proper fees, a license to a nonresident if the person meets the requirements of this Chapter or the person resides in a state that recognizes licenses issued by the Board. (2001-369, s. 1.)

#### § 74F-9. Fees.
The Board shall establish fees not exceeding the following amounts:

    (1) Issuance of a license $300.00
    (2) Renewal of a license $300.00
    (3) Examination $200.00
    (4) Reinstatement $250.00
    (5) Late fees $300.00
    (6) Apprentice license fee $300.00
    (7) Apprentice transfer fee $25.00.
(2001-369, s. 1; 2003-350, s. 7; 2013-370, s. 5.)

#### § 74F-10. Issuance, renewal, replacement, and transfer of licenses.
(a) The Board shall issue a license, upon payment of the license fee, to any applicant who
has satisfactorily met the requirements of this Chapter as administered by the Board. Licenses shall
show the full name of the person and an identification number and shall be signed by the chair and
one other officer of the Board.
(b) All licenses shall expire three years after the date they were issued unless renewed. All
applications for renewal shall be filed with the Board and shall be accompanied by the renewal fee
as required by G.S. 74F-9. A license that has expired for failure to renew may be reinstated after the
applicant pays the late and reinstatement fees as required by G.S. 74F-9. If an applicant whose
license has expired can show good cause to the Board the reason for allowing the license to expire,
the Board, in its discretion, may adjust the renewal and reinstatement fees accordingly.
(c) The Board shall replace any license that is lost, destroyed, or mutilated subject to rules
established by the Board.
(d) A license may not be transferred or assigned. (2001-369, s. 1; 2013-370, s. 6.)

#### § 74F-11. Photo identification.
Every person licensed under this Chapter shall be issued a photo identification card by the
Board. The card shall display a current photograph of the person, the person's name, address, and
telephone number. The licensee shall have the photograph identification card available for
inspection while performing locksmith services. (2001-369, s. 1.)

#### § 74F-12. Posting licenses; advertisements.
(a) Every locksmith issued a license under this Chapter shall display the license
prominently in the locksmith's place of business.
(b) All advertisements for locksmith services shall include a valid license number issued
by the Board. The license number of the owner of the locksmith company shall satisfy the
requirements of this subsection. (2001-369, s. 1; 2013-370, s. 7.)

#### § 74F-13. Responsibilities of employers.
Every licensee under this Chapter shall provide to the Board the names of each person
employed by the licensee who either performs locksmith services or has access to locksmith tools.
The licensee shall notify the Board within 30 days of any change in the information provided
pursuant to this section. (2001-369, s. 1.)

#### § 74F-14. Customer identification.
When opening a locked door to any vehicle or residential or commercial property, a licensee
shall make a reasonable effort to verify that the customer is the legal owner of the vehicle or
property or is authorized by the legal owner to gain access to the vehicle or property. (2001-369, s.
1.)

#### § 74F-15. Disciplinary procedures.
(a) The Board may deny or refuse to renew, suspend, or revoke a license or apprenticeship
designation if the licensee, apprentice, or applicant:
(1) Gives false information to or withholds information from the Board in
procuring or attempting to procure a license.
(2) Has been convicted of or pled guilty or no contest to any of the crimes listed in
G.S. 74F-18(a)(2).
(3) Has demonstrated gross negligence, incompetency, or misconduct in
performing locksmith services.
(4) Has willfully violated any of the provisions of this Chapter.
(b) The Board may assess the costs of disciplinary action, including attorneys' fees, against
an applicant or licensee found to be in violation of this Chapter or rules adopted by the Board.
(2001-369, s. 1; 2003-350, s. 8; 2013-370, s. 8.)

#### § 74F-16. Exemptions.
The provisions of this Chapter do not apply to:

(1) An employee of a licensed locksmith when acting under the direct control and
supervision of the licensed locksmith. For purposes of this subdivision, "direct
control and supervision" means that a licensed locksmith is required to
physically accompany the employee to the premises where locksmith services
are to be performed.
(1a) An employee of a locksmith company performing administrative duties only.
For purposes of this section, "administrative duties" means managing the daily
operations of an office in a locksmith company, including performing clerical
tasks, answering telephones, and greeting customers.
(2) A person working as an apprentice pursuant to G.S. 74F-7.1.
(3) A person or business required to be licensed or registered by the North Carolina
Alarm Systems Licensing Board pursuant to Chapter 74D of the General
Statutes, when acting within the scope and course of the alarm systems license
or registration.
(4) A person or business providing any of the following services so long as the
person or business does not represent himself, herself, or itself as a locksmith:
a. A towing service, or its employee, when providing services in the
normal course of its business.
b. An automotive repair business, or its employee, when opening a vehicle
to perform service on the vehicle.
c. A repossessing company, or its employee, while repossessing a vehicle.
d. A motor vehicle dealer as defined in G.S. 20-286(11), or a motor club as
defined in G.S. 58-69-1 when opening automotive locks in the normal
course of the dealer or club's business duties.
(5) A property owner, or the owner's employee, when providing locksmith services
on the property owner's property, so long as the owner or employee does not
represent himself or herself as a locksmith. For purposes of this section,
"property" means, but is not limited to, a hotel, motel, apartment, condominium,
commercial rental property, and residential rental property.
(6) A merchant, or retail or hardware store, so long as all of the following apply:
NC General Statutes - Chapter 74F 7
a. It is lawfully duplicating keys or installing, servicing, repairing,
rebuilding, reprograming, rekeying, or maintaining locks in the normal
course of its business.
b. It maintains a physical location in this State.
c. It maintains a sales and use tax permit in accordance with G.S.
105-164.16.
d. It does not represent itself as a locksmith.
(7) A member of a law enforcement agency, fire department, or other government
agency who, when acting within the scope and course of the member's
employment with the agency or department, opens locked doors to vehicles,
homes, or businesses.
(8) A salesperson while demonstrating the use of locksmith tools to persons
licensed under this Chapter.
(9) A general contractor licensed under Article 1 of Chapter 87 of the General
Statutes when acting within the scope and course of the general contractor
license, or an agent or subcontractor of a licensed general contractor when
acting within the ordinary course of business.
(10) A person or business when lawfully installing or maintaining a safety lock
device on a wastewater system when the safety lock device is required by
permit or requested by the owner of the wastewater system, provided the person
or business does not represent itself as a locksmith. For purposes of this
subdivision, "wastewater system" has the same meaning as in G.S. 130A-334.
(11) Any person or firm that sells gun safes or locking devices for firearms during
the course of the sale of gun safes or locking devices for firearms, so long as the
person or firm, or the firm's employee, does not represent himself, herself, or
itself as a locksmith.
(12) A person while performing a locksmith service in an emergency situation
without receiving any compensation for this service and who does not advertise
those services. (2001-369, s. 1; 2003-350, ss. 9, 10, 10.1; 2013-370, s. 9;
2014-120, s. 9.)

#### § 74F-17. Injunctions.
The Board may apply to the superior court for an order enjoining violations of this Chapter.
Upon a showing by the Board that any person has violated this Chapter, the court may grant
injunctive relief. (2001-369, s. 1.)

#### § 74F-18. Criminal history record checks of applicants for licensure or apprentice designation.

(a) Definitions. – The following definitions shall apply in this section:
(1) Applicant. – A person applying for licensure as a locksmith pursuant to
G.S. 74F-7 or apprentice designation pursuant to G.S. 74F-7.1.
(2) Criminal history. – A history of conviction of a state or federal crime, whether a
misdemeanor or felony, that bears on an applicant's fitness for licensure to
practice locksmithing. The crimes include the criminal offenses set forth in any
of the following Articles of Chapter 14 of the General Statutes: Article 5,
Counterfeiting and Issuing Monetary Substitutes; Article 5A, Endangering
NC General Statutes - Chapter 74F 8
Executive and Legislative Officers; Article 6, Homicide; Article 7B, Rape and
Other Sex Offenses; Article 8, Assaults; Article 10, Kidnapping and Abduction;
Article 13, Malicious Injury or Damage by Use of Explosive or Incendiary
Device or Material; Article 14, Burglary and Other Housebreakings; Article 15,
Arson and Other Burnings; Article 16, Larceny; Article 17, Robbery; Article
18, Embezzlement; Article 19, False Pretenses and Cheats; Article 19A,
Obtaining Property or Services by False or Fraudulent Use of Credit Device or
Other Means; Article 19B, Financial Transaction Card Crime Act; Article 20,
Frauds; Article 21, Forgery; Article 26, Offenses Against Public Morality and
Decency; Article 26A, Adult Establishments; Article 27, Prostitution; Article
28, Perjury; Article 29, Bribery; Article 31, Misconduct in Public Office;
Article 35, Offenses Against the Public Peace; Article 36A, Riots, Civil
Disorders, and Emergencies; Article 39, Protection of Minors; Article 40,
Protection of the Family; Article 59, Public Intoxication; and Article 60,
Computer-Related Crime. The crimes also include possession or sale of drugs
in violation of the North Carolina Controlled Substances Act in Article 5 of
Chapter 90 of the General Statutes and alcohol-related offenses including sale
to underage persons in violation of G.S. 18B-302 or driving while impaired in
violation of G.S. 20-138.1 through G.S. 20-138.5. In addition to the North
Carolina crimes listed in this subdivision, such crimes also include similar
crimes under federal law or under the laws of other states.
(b) All applicants for licensure or apprentice designation shall consent to a criminal history
record check. Refusal to consent to a criminal history record check may constitute grounds for the
Board to deny licensure or apprentice designation to an applicant. The Board shall ensure that the
State and national criminal history of an applicant is checked. The Board shall be responsible for
providing to the State Bureau of Investigation the fingerprints of the applicant to be checked, a
form signed by the applicant consenting to the criminal record check and the use of fingerprints and
other identifying information required by the State or National Repositories of Criminal Histories,
and any additional information required by the State Bureau of Investigation. The Board shall keep
all information obtained pursuant to this section confidential. The Board shall collect any fees
required by the State Bureau of Investigation and shall remit the fees to the State Bureau of
Investigation for expenses associated with conducting the criminal history record check.
(c) If an applicant's criminal history record check reveals one or more convictions listed
under subdivision (a)(2) of this section, the conviction shall not automatically bar licensure. The
Board shall consider all of the following factors regarding the conviction:
(1) The level of seriousness of the crime.
(2) The date of the crime.
(3) The age of the person at the time of the conviction.
(4) The circumstances surrounding the commission of the crime, if known.
(5) The nexus between the criminal conduct of the person and the job duties of the
position to be filled.
(6) The person's prison, jail, probation, parole, rehabilitation, and employment
records since the date the crime was committed.
(7) The subsequent commission by the person of a crime listed in subdivision (a)(2)
of this section.
NC General Statutes - Chapter 74F 9
If, after reviewing these factors, the Board determines that the applicant's criminal history
disqualifies the applicant for licensure, the Board may deny licensure or apprentice designation of
the applicant. The Board may disclose to the applicant information contained in the criminal
history record check that is relevant to the denial. The Board shall not provide a copy of the
criminal history record check to the applicant. The applicant shall have the right to appear before
the Board to appeal the Board's decision. However, an appearance before the full Board shall
constitute an exhaustion of administrative remedies in accordance with Chapter 150B of the
General Statutes.
(d) Limited Immunity. – The Board, its officers, and employees, acting in good faith and in
compliance with this section, shall be immune from civil liability for denying licensure or
apprentice designation to an applicant based on information provided in the applicant's criminal
history record check. (2003-350, s. 11; 2004-177, s. 2; 2012-12, s. 2(ee); 2015-181, s. 47;
2014-100, s. 17.1(o); 2015-181, s. 47; 2023-134, s. 19F.4(xx).)

## NC Locksmith Licensing Board

## Anticipated Expenditures